Economic Evaluation of Mis-speficied Geographically Weighted Regression Models for Site-specific Nitrogen Management

### 2023-06-24 Writing/manuscript_gwr.rmd
minor edit

**Abstract**: Geographically weighted regression (GWR) has been presented as a prominent tool to estimate site-specific yield response functions to derive site-specific input rate recommendations from data generated by on-farm precision experimentation (OFPE). However, the current application of GWR in this context assumes quadratic yield response function even though many of the yield-input relationships has been shown to be better represeted by the quadratic-plateau functional form. This is due to lack of readily available methods and satistical packages that allow researchers to estimate non-linear model undet the GWR framework, including quadratic-plateau. The key objective of this paper is to investigate the possibility of bias in estimating the economic value of site-specific recommendation maps based on GWR with quadratic functional form assumption when in reality the true underlying yield-input relationship is quadratic-plateau. We consider two sets of functional-form specifications used by two recent published studies of site-specific input management that apply GWR to data from OFPE. We show that both GWR approaches suffer significant positive bias, which can lead to large over-estimation of the economic value of applying GWR methods. We reveal that in many circumstances using GWR is far less profitable than claimed by previous studies, which provides a necessary cautionary message to researchers who might otherwise “over-sell” to other researchers the value of using GWR methods in empirical work involving the estimation of site-specific input management. We encourage the statistical community to extend the current GWR capability by allowing researchers to model yield-inpute response more flexibly than currently possible.

>>>>

**Abstract**: Geographically weighted regression (GWR) has been presented as a prominent tool to estimate site-specific yield response functions to derive site-specific input rate recommendations from data generated by on-farm precision experimentation (OFPE). However, the current application of GWR in this context assumes quadratic yield response function even though many of the yield-input relationships has been shown to be better represented by the quadratic-plateau functional form. This is due to lack of readily available methods and satistical packages that allow researchers to estimate non-linear model under the GWR framework, including quadratic-plateau. The key objective of this paper is to investigate the possibility of bias in estimating the economic value of site-specific recommendation maps based on GWR with quadratic functional form assumption when in reality the true underlying yield-input relationship is quadratic-plateau. We consider two sets of functional-form specifications used by two recent published studies of site-specific input management that apply GWR to data from OFPE. We show that both GWR approaches suffer significant positive bias, which can lead to large over-estimation of the economic value of applying GWR methods. We reveal that in many circumstances using GWR is far less profitable than claimed by previous studies, which provides a necessary cautionary message to researchers who might otherwise “over-sell” to other researchers the value of using GWR methods in empirical work involving the estimation of site-specific input management. We encourage the statistical community to extend the current GWR capability by allowing researchers to model yield-inpute response more flexibly than currently possible.
